Facility and Time Step Discretization Effects in Fuel Cycle Simulation

### 2015-09-30 paper.tex
More puinv plot discussion. make some figs wider.

\begin{figure*}[!h]
    \centering
    \includegraphics[width=0.8\textwidth]{exp2/power.eps}
    \caption[Generated power]{
        Generated power for all four cases.  A fuel shortage occurs from about
        year 100 to about year 140.  In cases MI and QI with individual reactor
        modeling, refueling cycles become much more synchronized during and
        after the fuel shortage resuting in large power fluctuations - an artifact of the
        \emph{cycle staggering effect}.
    }
    \label{fig:power}
\end{figure*}

\begin{figure*}[!h]
    \centering
    \includegraphics[width=0.8\textwidth]{exp2/power-rel.eps}
    \caption[Normalized power]{
        Generated power is normalized to the expected 1\% exponential growth
        curve for all four cases.  A fuel shortage occurs from about year 100
        to about year 140.  The \emph{cycle staggering effect} can be seen as
        excessive divergence above and below $1.0$ in cases MI and QI where
        the shortage increases cycle synchronization significantly. The
        \emph{quantized shutdown effect} can also be seen in the discrepency
        between quarterly and monthly time steps for fleet reactor modeling
        (cases MF and QF) during the shortage; a longer time step causes more
        reactor outage than necessary. The consistent jumping up and down for
        the fleet reactor cases (even pre-shortage) is due to the mis-alignment of reactors going
        offline (multiples of 12 months) and the build period (21 months).
    }
    \label{fig:power-rel}
\end{figure*}

@@

The large differences seen between monthly and quarterly time step cases in
Figures \ref{fig:unfueled} and \ref{fig:badshare} are primarily caused by the
\emph{inventory drawdown effect}.  Figure \ref{fig:puinv} helps to visualize
this effect.  The black dots in the figure represent impulse flows of
available Pu out of the fuel fab and into fast reactors.  Blue lines represent
the flows into Pu inventory available for making fuel. The in flows are not
shown for the individual reactor cases because they are somewhat messy and
obscure other information in the plots. The other colored curves show Pu inventory
available for making fast reactor fuel.  When the inventory curves meet and
force down the out-flows, fuel shortages occur.  While this is most clear for
cases MF and QF, the individual reactor scenarios (cases MI and QI) also have many
points during the shortage where the outflow point lies exactly on top of the
inventory point - indicating that all available inventory is being
transferred.

@@

Throughout most of the shortage, the aggregate outflow of Pu from separated
inventory is roughly the same in cases MF and QF.  However, due to the longer
time steps in case QF, the shortage starts slightly sooner because Pu inventory
must be maintained at a higher limit as described above.  The small difference
in outage start time has feed-back effects.  Not only are extra reactors
offline, but those extra reactors are also discharging less aggregate spent
fuel for recycling.  This farther increases the discrepency between the two
cases. Eventually the shortage in case MF ends slightly before it does in case
QF. This effect feedback occurs because the reactors in question have a
conversion ratio greater than 1.0.

The striations visible in the monthly flows for case MI in Figure
\ref{fig:puinv} are a characteristic of the synchronization of reactor
refueling caused by the shortage.  They begin at the end of the fuel shortage
window when many reactors come online together and are resonances that the
inventory level jumps between.  These resonances gradually disappear as the
simulation progresses farther in time beyond the shortage and refueling cycles
become more staggered.

@@

\begin{figure*}[!h]
    \centering
    \includegraphics[width=0.8\textwidth]{exp2/puinv-compare.eps}
    \caption{
        Separated Pu inventories for all four cases.  Differences in modeling
        choices are not particularly significant until about year 100 when
        fuel shortages begin.  These differences gradually decay away starting
        in year 140 after the fuel shortages have ebbed.  Higher aggregate Pu
        inventory  during the shortage is correlated with a longer during and
        post shortage recovery period.
    }
    \label{fig:puinv-compare}
\end{figure*}

>>>>

\begin{figure*}[!h]
    \centering
    \includegraphics[width=1.0\textwidth]{exp2/power.eps}
    \caption[Generated power]{
        Generated power for all four cases.  A fuel shortage occurs from about
        year 100 to about year 140.  In cases MI and QI with individual reactor
        modeling, refueling cycles become much more synchronized during and
        after the fuel shortage resuting in large power fluctuations - an artifact of the
        \emph{cycle staggering effect}.
    }
    \label{fig:power}
\end{figure*}

\begin{figure*}[!h]
    \centering
    \includegraphics[width=1.0\textwidth]{exp2/power-rel.eps}
    \caption[Normalized power]{
        Generated power is normalized to the expected 1\% exponential growth
        curve for all four cases.  A fuel shortage occurs from about year 100
        to about year 140.  The \emph{cycle staggering effect} can be seen as
        excessive divergence above and below $1.0$ in cases MI and QI where
        the shortage increases cycle synchronization significantly. The
        \emph{quantized shutdown effect} can also be seen in the discrepency
        between quarterly and monthly time steps for fleet reactor modeling
        (cases MF and QF) during the shortage; a longer time step causes more
        reactor outage than necessary. The consistent jumping up and down for
        the fleet reactor cases (even pre-shortage) is due to the mis-alignment of reactors going
        offline (multiples of 12 months) and the build period (21 months).
    }
    \label{fig:power-rel}
\end{figure*}

@@

The large differences seen between monthly and quarterly time step cases in
Figures \ref{fig:unfueled} and \ref{fig:badshare} are primarily caused by the
\emph{inventory drawdown effect}.  Figure \ref{fig:puinv} helps to visualize
this effect.  The black dots in the figure represent impulse flows of
available Pu out of the fuel fab and into fast reactors.  Blue lines represent
the flows into Pu inventory available for making fuel. The in-flows are not
shown for the individual reactor cases because they are somewhat messy and
obscure other information in the plots. The other colored curves show Pu inventory
available for fabricating fast reactor fuel.  When the inventory curves meet and
force down the out-flows, fuel shortages occur.  While this is most clear for
cases MF and QF, the individual reactor scenarios (cases MI and QI) also have many
points during the shortage where the outflow point lies exactly on top of the
inventory point - indicating that all available inventory is being
transferred.

@@

Throughout most of the shortage, the aggregate outflow of Pu from separated
inventory is roughly the same in cases MF and QF.  However, due to the longer
time steps in case QF, the shortage starts slightly sooner because Pu inventory
must be maintained at a higher limit as described above.  The small difference
in outage start time has feed-back effects.  Not only are extra reactors
offline, but those extra reactors are also discharging less aggregate spent
fuel for recycling.  This farther increases the discrepency between the two
cases. Eventually the shortage in case MF ends slightly before it does in case
QF. This feedback effect occurs because the reactors in question have a
conversion ratio greater than 1.0.

The horizontal striations visible in the monthly flows for case MI in Figure
\ref{fig:puinv} are a characteristic of the synchronization of reactor
refueling caused by the shortage.  They begin at the end of the fuel shortage
window when many reactors come online together and are resonances that the
inventory level jumps between.  These resonances gradually disappear as the
simulation progresses farther in time beyond the shortage and refueling cycles
become more staggered.

The in-flows shown for cases MF and QF show a periodic pulsing that begins
during the shortage in about year 120.  This pusling actually has nothing to
do with the shortage and is caused by the retirement of fast reactors.  The
first fast reactors are deployed in year 35 and have an 80 year lifetime.
When they retire, they discharge an entire core's worth of fuel rather than a
single batch.  After a 7-year cooling period, the first discharged full cores
begin to make it through the recycle loop in year 122.  Another interesting
feature visible in all four cases in Figure \ref{fig:puinv} is the minority
fraction of out-flow points that are higher than the bulk majority. These
higher out-flows occur once every 21 months when new reactors are built. The
new reactors require a full core of fresh fuel rather than the single batch
needed when just refueling.

@@

\begin{figure*}[!h]
    \centering
    \includegraphics[width=1.0\textwidth]{exp2/puinv-compare.eps}
    \caption{
        Separated Pu inventories for all four cases.  Differences in modeling
        choices are not particularly significant until about year 100 when
        fuel shortages begin.  These differences gradually decay away starting
        in year 140 after the fuel shortages have ebbed.  Higher aggregate Pu
        inventory  during the shortage is correlated with a longer during and
        post shortage recovery period.
    }
    \label{fig:puinv-compare}
\end{figure*}
